---
title: "Using `eesim`"
author: "Sarah Koehler and Brooke Anderson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(eesim)
library(ggplot2)
```

## Overview of package

This package allows you to simulate time series of environmental health data, to allow you to perform simulation-based power analyses and evaluate model performance.

## Basic example of using the package

You can use the eesim function to generate data, fit models, and assess the models:

```{r}
eesim(n_reps = 1000, n = 1000, central = 100, sd = 10, exposure_type = "continuous", exposure_trend = "cos1", exposure_amp = .6, average_outcome = 22, rr = 1.01)
```


## Comparing two models

As a basic example, you can use this package to simulate a continuous environmental exposure with seasonal and year-round trends and a count outcome. You can then fit a case-crossover and GLM model to this data and compare performance.


## More detailed examples
